Title here

### 2017-01-10 portfolio/text/ms.Rmd

<!-- Introduction, Methods, Results, Discussion, Conclusion -->
<!-- 600, 900, 500, 800, 200, -->

@@

<!-- TODO < 250 words -->

Individuals relying on the extraction of natural resources for their livelihood
--- fishers, farmers, miners, and loggers, among others --- face high levels of
income variability driven by a mix of environmental, biological, management, and
economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels fishing a greater diversity of species have less revenue
variability than those fishing a lower diversity. However, it is unclear if
these benefits extend to the actions of individual fishers and how year-to-year
changes in diversification affect revenue and revenue variability. Here, we
identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities. We show that,
despite increasing specialization over the last 30 years, fishing permits with
higher species diversity reduces individual revenue variability and fishing an
additional permit is associated with higher revenue and lower variability.
However, increasing species diversity within the constraints of existing permits
has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
probability) associated with increased revenue uncertainty the following year.
Our results demonstrate that the most effective option for individuals to
decrease revenue variability is to purchase permits to participate in additional
or more diverse fisheries. However, this option is expensive, often limited by
regulations such as catch share programs, and consequently unavailable to many
individuals. With increasing environmental variability in a changing climate, it
will be particularly important that individuals relying on natural resources for
their livelihood have effective strategies to reduce income variability.

<!-- TODO: take home is the management context is key and controls show that
it's more complicated. TODO mentioned that specializing tends to be profitable
-->

@@

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, external forces may restrict the available
options to dampen income variability. Otherwise well-intentioned regulations may
limit how individuals diversify, or incentivize against diversification. For
example, in terrestrial agriculture, crop subsidies in the U.S. may incentivize
some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, external market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g. counties or communities; [@kasperski2013;
@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
between diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
countries have passed --- fishers must own a license to participate in virtually
all fisheries [@hilborn2003a]. While such regulations have helped reduce
overfishing, they limit the available strategies to reduce income variability.
For example, many permits limit fishing to a single species, diversifying by
purchasing additional fishing permits can be expensive[^1], and some fisheries
limit the entry of new participants through individual fishing quotas (IFQs)
[@grimm2012].

[^1]: For example, it cost ~\$192K USD for a king and tanner crab permit in
Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.

Using a unique dataset collected over 30 years on individual fishers, rather
than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability. We develop a variance-function regression model
[@davidian1987; @hall1989], to simultaneously model changes in the mean and
variance of individual revenue as a function of covariates. We simultaneously
estimate the effect of species diversity (the inverse of Simpson's diversity
index [@simpson1949; @jost2006]) on individual revenue and revenue variability,
and estimate effects both within and across fishing strategies while controlling
for fishing effort and latent time effects (e.g. changing environmental
conditions). To our knowledge, this is the first application of this type of
model in any natural resource field.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold a single permit, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple-permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of this revenue declined somewhat from 1980--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
by an increase in salmon specialists in the 1980s and 1990s --- the trend
disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

Individual revenue variability decreases exponentially with higher mean fished
species diversity across combinations of state and federal fishing permits
("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
one unit higher mean species diversity had on average `r h1[[2]]`% (`r
h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern looks
qualitatively similar with or without individuals targeting salmon. As examples,
variability was highest for the herring roe gillnet and king crab fisheries.
Most strategies with the lowest estimated revenue variability combined halibut
with a salmon permit and had high levels of fished species diversity (Fig.
\ref{fig:across}).

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense.

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given set of permits is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
annual average revenue for fishers with the same set of permits, specializing
tends to be associated with increased revenue but also increased variability
(Fig. \ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}).
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a set of permits tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects involving halibut and sablefish permits to vary
separately before and after the introduction of IFQs. The introduction of IFQs
reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with only
halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
@edwards2004]. However, in practice, the ability of individuals to effectively
diversify is limited by regulation and market forces. Here we document the
effect of individual-level changes in species diversity on revenue and revenue
variability across three scales. We find that (1) individuals with sets of
permits that allow high levels of species diversity tend to have lower revenue
variability. (2) Adding a fisheries permit usually increases revenue and
decreases revenue variability. However, (3) within the constraints of a permit
and accounting for effort, specializing tends to increase revenue while both
specializing and generalizing tend to correlate with greater revenue variability
with considerable differences amongst fisheries. Thus, the most effective
mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is expensive,
limited by regulations, and consequently unavailable to many individuals.

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
findings within strategies, with respect to individual year-to-year behaviour,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, local knowledge and experience are
a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
fishing patterns may sometimes be associated with a learning or adaptation phase
as individuals become accustomed to new locations, gear, or species.
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers were forced to target other species or landed
more bycatch, adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our large-scale model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behaviour [@hilborn1985; @hilborn1985a;
@branch2006] and therefore not included in our large-scale model, that might
make a specific individual's revenue less predictable if species diversity
changes.

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial expense. Related barriers are apparent in
other natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008]. Another major barrier to
diversification can be management restrictions. In commercial fisheries,
regulations that you have been shown to reduce overfishing --- including
improved assessments [@melnychuk2017] and individual quota systems
[@costello2008] --- may ultimately restrict individual diversification
[@smith1986; @kasperski2013] or at least make developing a diverse fishing
portfolio more expensive. However, individual quota systems, in particular, may
also reduce income variability (Fig. S\ref{fig:ifq}A) perhaps offsetting the
effects of reduced diversification [@holland2016].

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. We expect future research examining diversification over
finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [@anderson2016].

With increasing environmental variability and extreme events in a changing
climate [@easterling2000], it is particularly important that individuals who
rely on natural resources as a primary source of income, such as fishers,
farmers, and forestry workers, have effective strategies to reduce risk. While
diversification can provide a powerful form of risk buffering, we have shown
that the relationship between individual changes to diversification and income
variability is complex and context specific. Furthermore, external market forces
can interact with management regulations, which have other considerable
benefits, to restrict individual diversification opportunities and promote
specialization. Although variability in individual revenue may not be actively
incorporated into management decisions, analyses such as ours help to understand
the range of diversification strategies available, trade-offs inherent in such
management decisions, and how these trade-offs may change over time.

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
sethi2014], we calculated effective species diversity for permit holder $i$ and
time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
[@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index [@hirschman1964]) weighted by revenue $R$:

>>>>

<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
<!-- 600, 900, 500, 800, 200, -->

@@

<!-- TODO slash length of abstract -->

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability. Fishers, loggers, farmers,
and miners, among others, face high levels of income uncertainty driven by a mix
of environmental, biological, and economic variability. A key component
underlying the sustainability of these industries is identifying both how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that targeting a diversity of
species can reduce income variability. However, any diversity-variability
relationship is embedded in a regulatory framework that restricts individual
ability to change fishing behaviour. Here we identify three primary axes by
which fishers in Alaska, one of the world's most productive fishery regions, can
potentially diversify fishing activities and investigate how individual actions
along each diversity axis affect both revenue and revenue variability. We
construct a novel statistical model to simultaneously assess the value to
individual fishers of choosing to (1) own permits that allow targeting a diverse
fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
a greater species diversity within the constraints of a given permit. We show
contrasting effects of diversity on revenue and revenue variability along the
three axes. First, we show that fishing permits with higher species diversity
reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
per unit increase in mean effective species diversity). Next, we show that
fishing an additional permit is associated with higher revenue and lower
variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
IQR] decline in standard deviation for fishers with halibut and another permit
vs. only a halibut permit). Finally, we show that increasing species diversity
within the constraints of a permit set has a fishery-dependent effect on revenue
and is usually (`r g2_dens`% probability) associated with increased uncertainty
in revenue the following year. Thus, while there are clear benefits to some axes
of diversification, others are less effective. Importantly, the most effective
mechanism for increasing revenue and decreasing risk --- purchasing additional
permits to participate in additional fisheries --- was also the most costly.
Furthermore, the least costly option of increasing catch diversity within an
existing permit was associated with increased revenue variability. Thus, over
the past 30 years as the Alaskan commercial fishing fleet has become
increasingly specialized and most individuals only participate under a single
permit, the practical benefits accrued from increased species diversification
are limited.

Our results suggest that the benefit of species diversity for revenue
variability is mostly an across-permit effect, not a benefit that can be easily
harnessed by an individual permit holder. Effective methods of diversification
--- purchasing additional permits or switching permits --- are costly and often
limited by regulations such as limited entry programs and individual
transferable quotas (ITQs). Take home ...

@@

<!-- Papers to cite [@smith1986] -->

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high year-to-year
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability. For example, drought and flooding are a major source of risk for
agricultural food security and farmers' incomes [@gregory2005], and catastrophic
disease outbreaks and wildfires increase risk for the logging industry
[@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, external forces may restrict the available
options to dampen income variability. Otherwise well-intentioned regulations may
limit how individuals diversify, or incentivize against diversification. For
example, in terrestrial agriculture, crop subsidies in the US may incentivize
some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, external market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
the effects of diversification on revenue do so at aggregate levels that combine
individuals (e.g. households, counties, or communities; [@kasperski2013;
@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
between diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. Like much of the rest of the world, the days of open,
unregulated fisheries in most developed countries have passed --- fishers must
own a license to participate in virtually all fisheries [@hilborn2003a]. And
while such regulations have helped reduce overfishing, they limit the available
strategies to reduce income variability. For example, many permits limit fishing
to a single species, diversifying by purchasing additional fishing permits can
be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) [@grimm2012].

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fished species
diversity [OLE: Call this diversity of catch rather than fished?] on individual 
revenue and revenue variability that builds on results
from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.

We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold single permits, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean effective
species diversity of this revenue declined somewhat from 1980--2000 but has
since remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
driven by an increase in salmon specialists in the 1980s and 1990s --- the trend
disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

Individual revenue variability decreases exponentially with higher mean fished
species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
effort, strategies with one unit higher mean effective species diversity had on
average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
pattern looks qualitatively similar with or without individuals targeting
salmon. As examples, variability was highest for the herring roe gillnet and
king crab fisheries. Most strategies with the lowest estimated revenue
variability combined halibut with a salmon permit and had high levels of fished
species diversity (Fig. \ref{fig:across}).

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016 REF;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies, `r g2_dens`% of the
probability density supports a positive relationship between specialization and
revenue.

<!--This benefit of specializing is most apparent after accounting for fishing
effort ().-->

However, changing (increasing or decreasing) an individual's fished species
diversity within the constraints of a permit set tends to be associated with
greater expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects to vary separately before and after the introduction of
IFQs for strategies involving halibut and sablefish permits. The introduction of
IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies with a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, the ability of individuals to
effectively diversify is limited by regulation and market forces in practice.
Here we document the effect of changing species diversity on revenue and revenue
variability across three scale. Each scale reflects individual-level choices
that can affect species diversity. We find that individuals with
fishing permit sets that allow high levels of species diversity tend to have
lower revenue variability. Adding a fisheries permit usually
increases revenue and decreases revenue variability. Within the
constraints of a permit and accounting for effort, specializing (reducing
species diversity) tends to increase revenue while both specializing or generalizing
tends to be associated with greater revenue variability - though there are considerable
differences among fisheries. Thus, the most effective mechanism by which
species diversity can lower revenue variability for individual fishers ---
switching or adding fishing permits --- is costly, limited by regulations,
and consequently unavailable to many individuals in practice.

Our results shows additional support for the idea that individual fishers can
reduce their income variability by diversifying their catch portfolios. However,
most individuals do not appear to change their catch diversity substantially
from year to year [OLE: is there a supplemental figure or statistic that speaks to this point?], and most strategies adopted by Alaskan fishermen lie in
a high-risk, high-reward quadrant in which reducing diversification increases
both revenue variability and expected revenue. Previous work has suggested that
changes to fisheries management such as limited entry or catch share programs
may inadvertently act as barrier to diversification and promote specializing
(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
specialized fishing strategies may be associated with increased revenue
variability, the benefit may be that they are also associated with increased
revenue.

External market forces may interact in concert with management
regulations to promote specialization for fishers or other individuals dependent
on natural resources. In terrestrial agriculture for instance, there are 
many instances of specializing being profitable; drivers include the ability to
specialize on high-profit cash crops (tobacco, cocoa), the availability of
government agricultural subsidies (soybeans, corn in the USA), and farming on
a large industrial scale (REF). With increasing environmental variability and extreme
events in a changing climate (REF), it is particularly important that
individuals who rely on natural resources as a primary source of income have
strategies to reduce risk. Though variability in individual revenues may not be
actively incorporated into many management decisions, analyses like ours help to
understand the range of diversification strategies available and how their
benefits may change over time.

Worldwide, fisheries confront the challenge of sustainable management (). While many fisheries are well managed and productive, many are also rebuilding from periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid in fisheries rebuilding is to improve fishing gear modify regulations to reduce the unintended harvest rebuilding stocks while allowing the continued harvest of healthy stocks (need REF). Such bycatch reduction is an important goal for both ecological (Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks 2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory changes for bycatch reduction as well as other reasons gradually lead to fisheries that target one or a few species and individual fishers wishing to participate in multiple fisheries must own (or purchase) multiple permits. Thus a general consequence of fisheries regulations that are shown to improve sustainability of fish stocks and avoid overfishing - including improved assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et al. 2008 ) - may ultimately restrict individual diversification or at least make developing a diverse fishing portfolio more costly. Add Alaska example of IFQs...?

Chakravorty, U. and Nemoto, K. (2001) Modeling the effects of area closure and tax policies: a spatial-temporal model of the hawaii longline fishery. Marine Resource Economics 15, 179–204.

Costello C, Gaines SD, Lynham J (2008) Can catch shares prevent fisheries collapse? Science 321(5896):1678–1681

Crowder, L.B. and Murawski, S.A. (1998) Fisheries bycatch: implications for management. Fisheries Management 23, 8–17.

Curtis, R. and Hicks, R.L. (2000) The Cost of sea turtle preservation: the case of Hawaii’s Pelagic longliners. American Journal of Agricultural Economics 82, 1191– 1197.

Hilborn R, Ovando D. 2014 Reflections on the success of traditional fisheries management. ICES J. Mar. Sci. 71, 1040–1046.

Lewison, R.L., Crowder, L.B., Read, A.J. and Freeman, S.A. (2004) Understanding impacts of fisheries bycatch on marine megafauna. Trends in Ecology & Evolution 19, 598–604.

Read, A.J., Drinker, P. and Northridge, S. (2006) Bycatch of marine mammals in US and global fisheries. Conservation Biology 20, 163–169.

### Reconciling the effects at different levels of species diversity

- why variability/uncertainty might increase with changes in diversity...

- and how fit in with @kasperski2013, @sethi2012b, @sethi2014

- part of the difference is individuals vs vessels. E.g. @kasperski2013
  suggested that diversity declined following IFQs for halibut sablefish
  vessels. The reason for the decline was that the vessels were allowed to fish
  halibut for more days (because people were sharing boats). At the level of
  people, we wouldn't expect the diversity of someone who only had
  halibut/sablefish permit to be affected by IFQs. The person, not the vessel
  holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq

- year to year risky... But maybe evens out/better over time

- diversity within permit - often bycatch, limited room for intentional
  diversification (unless you have a M or salmon permit)

- risk results not really risk, but unpredictability given our covariates ...
  don't change anything and it's most predictable how well you'll do. What's
  unpredictable in our model, might be more predictable on a year to year,
  species by species basis for individual permit holder

- CV vs year-to-year effect for one person

Other:

- specializing can be profitable within some fisheries. Eric: analog here in
  farming literature is that diversification for small scale farmers generally
  good, except when they specialize on high demand/high value crops (tobacco,
  cocoa). Seems like benefits of diversification are also smaller for
  large-scale industrial farmers.

### Limitations + obvious extensions paragraph by Ole:

Our analysis provides a synthetic view of the effects of species diversity on individuals over 30 years in one of the most productive fisheries regions in the world. However, our analysis simultaneously uses data from across the broad geography of Alaska and over 30 years, potentially obscuring regional or time-variation in the consequences of diversification on revenue and revenue variability. We expect future research examining diversification over finer spatial and temporal scales will uncover additional interesting relationships. Furthermore, in addressing the overall pattern, we have averaged out several important of major discrete events (e.g., the Exxon Valdez oil spill in 1989) that had significant consequences for the ecosystem and social system in Alaska (general EVOS consequences REF). We do not wish to minimize the important of such events, but note that it is important to understand the benefits of diversification in the context of rare and extreme disturbances which occur in marine systems worldwide.

There are a number of limitations to our model and extensions that could further
our understanding of the role of diversification...

- could vary by time and vary by mean diversity

- spatial/geographic variation

- nonlinear effects

- response to specific events like EVOS

### Implications for limited entry, ITQs, and catch shares

- concerning because in many areas of the world and fisheries, regulations and
  market forces have encouraged specialization / limited (effective)
  diversification (e.g. multiple permits or diverse permits)

- locking into permits limits ability to pick lower risk strategies and is
  getting people locked into low diversity permits that may be doing well now,
  but not necessarily always

### Conclusions + recommendations

- bring it out to wider than fisheries (farming, timber industry, ...)

- 2 dimensions to think about (revenue and variability); and the trade-off is
  fishery specific: effects within a permit strategy are permit specific;
  sometimes make more money by specializing... some highly variable single
  species permits make a ton of money

- broadly, diversification can provide a powerful form of protection from
  multiple sources of risk to income for those that rely on extracting natural
  resources, but not all forms of diversification are equal, and on an
  individual level there can be economic benefits to specializing and risks to
  adjusting ones strategy

- with increasing environmental variability and extremes due to climate change
  those that rely on natural resources for their livelihoods need approaches to
  buffer risk more than ever

- regulations that limit the ability to diversify can have other considerable
  benefits but it's important to be aware of trade-offs inherent in these
  regulations when making management decisions

We fit a hierarchical variance-function regression model to 30 years of
commercial fisheries revenue data in Alaska to simultaneously estimate the
effect of changes in species diversity on individual's revenue and revenue
variability.

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@kasperski2013], we calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:
